Fluid-mechanical erosion creates anisotropic porous media

### 2022-05-06 docs/PhysicsPaper/StokesErosion.tex
NM

@@

>>>>

% From Nick's Latex Notes
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand {\bq} {\bvec{q}}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\qavg}{\bar{q}}
\newcommand{\pavg}{\bar{p}}

@@

%-----------------------------------------------------------------------------------------------%
\section{Measuring permeability and other porous-media properties}
\label{sec:}
%-----------------------------------------------------------------------------------------------%
Darcy's law is given by
\begin{equation}
\bq = - \bvec{K} \grad \phi
\end{equation}
where $\bq$ is the specific discharge, $\phi = z + p/\gamma$ is the hydraulic head, and $\gamma = \rho g$. The rank-2 tensor $\bvec{K}$ is the {\em conductivity}, which is related to the {\em permeability} $\bvec{k}$ via $\bvec{K} = \bvec{k} \gamma/\mu$, where $\mu$ is the viscosity of the fluid. The units of conductivity are length-per-time (i.e.~velocity) and the units of permeability are length squared.
Perhaps the most standard parameter to characterize the porous medium is the permeability. We eliminate gravity from consideration to get
\begin{equation}
\bq = - \bvec{k} \grad p / \mu
\end{equation}
Henceforth, we set $\mu = 1$ by the non-dimensionalization.

Consider first an isotropic medium with {\em scalar} permeability $k$. Equivalently, we can introduce a {\em resistivity} simply defined as the reciprocal of permeability, $r = 1/k$. Then taking the horizontal component gives
\begin{equation}
\label{q1}
q_1(x,y) = -\frac{1}{r} \pderiv{p}{x}(x,y)
\end{equation}
Consider the horizontal discharge, $q_1(x,y)$, averaged over a vertical slice
\begin{equation}
\qavg := \frac{1}{2} \int_{-1}^{1} q_1(x_0, y) dy
\end{equation}
By conservation of mass, $\qavg$ is independent of the horizontal position $x_0$. Since $\bq$ is a volume-average of the tracer velocity, $\qavg$ can be equivalently obtained by averaging the tracer velocity, which is how we will measure this quantity in the simulations.

Next, consider the pressure averaged over a vertical slice
\begin{equation}
\pavg(x_0) := \frac{1}{2} \int_{-1}^{1} p(x_0, y) dy
\end{equation}
Integrating \eqref{q1} over the domain $[-x_0, x_0] \times [-1,1]$, applying the fundamental theorem of calculus, and rearranging gives
\begin{equation}
r = \left( \frac{\pavg(-x_0) - \pavg(x_0)}{2 x_0 \qavg} \right)
\end{equation}
This formula gives the resistivity based on measurements of the pressure and discharge.
